EoS-insensitive relations after GW170817

### 2019-01-27 paper/bin-love-after-gw170817.tex
more updates on the paper, as well as addressing Nicos comments

{\ny{Just some general comments so that we don't forget:
\begin{itemize}
\item The EoS-insensitive relations are called ``binary Love'', ``I-Love-Q'', ``C-Love'' and ``R-Love.'' There is no need to call them ``binary Love EoS-insensitive relations.'' That's too verbose and not what they are called in the literature. I am ok with describing these relations as EoS-insensitive instead of approximately universal, but let's not change the names. 
\item There is some serious tidying up that needs to be done to all of the text. Specifically, there are missing prepositions, like ``the'', in lots of places that makes the reading a bit choppy and we can improve. Zack, when you have a chance, please do a detail read through and fix the flow as much as you can. 
\item Eventually, we need to discuss the abstract and the intro in more detail. 
\item Let's not forget to check the acknowledgements for funding sources. 
\item Let's do a spell check. 
\item Let's remove all commented out text.  
\end{itemize}
}}
The dependence of pressure on density for cold supranuclear matter found inside neutron stars (NSs) remains to be one of the largest mysteries in both nuclear physics and astrophysics.
This internal structure of such stars, known as the equation of state (EoS), is exceedingly important, as it is the determining factor for many NS observables - such as the mass, the radius, and many more.
Unfortunately, terrestrial experiments can only study the EoS to around the nuclear saturation density ($\rho_{\text{sat}} \approx 2.5 \times 10^{14} \text{ g/cm}^3$)~\cite{Li:HeavyIon,Tsang:SymmetryEnergy,Centelles:NeutronSkin,Li:CrossSections,Chen:SymEnergy} (although some temperature-dependent heavy-ion collision experiments may probe higher densities \zc{Katerina do you know of any papers on this?} {\ny{Look up the PREX and CREX experiments at JLab (a good author is Horowitz, e.g. Abrahamyan:2012gp, Horowitz:2012tj, Horowitz:2013wha)}}), making NSs ideal laboratories for constraining ultra dense nuclear matter.

@@

%Universal relations
Unfortunately, current detectors are only sensitive enough to accurately measure the dominant tidal parameter, $\lambda_0$.
Previous work by Yagi and Yunes~\cite{Yagi:binLove} resolved this issue by finding approximately EoS-insensitive ``binary Love universal relations" between symmetric and anti-symmetric combinations of tidal deformabilities $\Lambda_{s,a}=\frac{1}{2}(\Lambda_1 \pm \Lambda_2)$.
This allows one to further break degeneracies between tidal parameters, resulting in both (i) a decrease in uncertainty upon parameter extraction of $\Lambda_s$ on the removal of parameter $\Lambda_a$ from the model, and (ii) the ability to measure the secondary tidal parameter upon measurement of the first~\cite{Katerina:residuals}.  
Use of these relations leads to up to an order of magnitude improvement in parameter estimation of $\Lambda_s$ through a simple Fisher analysis~\cite{Yagi:binLove}, also confirmed with the detailed analysis of~\cite{Katerina:residuals}.
Similar ``I-Love-Q" and ``C-Love" relations~\cite{Yagi:ILQ} have been found between individual NS observables: quadrupole moment , moment of inertia, tidal deformability, and compactness; similarly allowing the removal of tidal parameters from the model list, and allowing an automatic determination of one observable through the measurement of another.
However, these EoS-insensitive relations have been tuned to a collection of EoS models, and each have intrinsic uncertainties associated with their use.

%How we improved this
In this analysis, we aim to improve this important work by imposing the restrictions found on the EoS~\cite{LIGO:posterior} derived from the recent binary NS merger observation, GW170817~\cite{TheLIGOScientific:2017qsa}.
By restricting to only EoSs which lie within the the 90\% credible region on pressure as a function of density found in~\cite{LIGO:posterior}, we repeat the analyses done in~\cite{Yagi:binLove,Yagi:ILQ} and show an increase in universality for I-Love-Q, C-Love, and binary Love relations.
We also investigate the relationship between the NS radius and its tidal deformability, or the R-Love relations, determining if the uncertainties can be reduced.
In addition, we consider hybrid star EoSs seen in~\cite{Paschalidis2018}, which experience strong first-order transitions from hadronic matter to quark matter.
These provide a departure from the nuclear matter relations considered here - providing valuable insight into them.
Further, we estimate the viability of using improved EoS-insensitive relations on future GW detections by performing simple Fisher analyses to approximate when the statistical errors from parameter extraction become comparable to the systematic errors due to EoS variation in EoS-insensitive relations.
To estimate the effect of waveform mismodeling, we additionally consider two different tidal corrections to the injected PhenomD~\cite{PhenomDI,PhenomDII} waveform: 6PN~\cite{Wade:tidalCorrections} and NRTidal~\cite{Samajdar:NRTidal}

\subsection{Executive Summary}
\zc{Me and Kent discussed whether or not we should include this section or not, and we decided like having the quick summary for busy readers - especially with the increasing number of figs and tables (8 tables, 14 figures so far). We can discuss this further if anyone doesn't want it included :)}
{\ny{I'm ok with keeping this, but let's get the science finished first.}}
%What we did
In this paper we attempt to improve I-Love-Q, C-Love, and binary Love relations by imposing constraints on the EoS found by GW170817~\cite{LIGO:posterior,TheLIGOScientific:2017qsa}.
First, we generate two large samples of spectral EoSs~\cite{Lindblom:2018rfr}, where we (i) impose the restriction that they must be drawn from GW170817s 90\% credible posterior in pressure as a function of density, and (ii) don't impose any restrictions for comparison purposes.
Next, we follow the important works of~\cite{Yagi:binLove} and~\cite{Yagi:ILQ} to show how the ``constrained" set of EoSs show improved universality (binary Love, I-Love-Q, and C-Love) from both the previous works, and the ``unconstrained" sample, shown by Fig.~\ref{fig:ILQ}.
In addition, we determine the degree to which hybrid star EoSs obey binary Love relations.
Finally, through a simple Fisher analysis, we consider the value in improving EoS-insensitive relations by approximating when the statistical errors from parameter extraction become comparable to the systematic errors due to EoS variation in EoS-insensitive relations.

%Final Results
Upon use of the constrained set of EoSs, we find an improvement in binary Love universality by a factor of $\sim 59$\% for stars with mass ratios of $0.75$, in C-Love universality by a factor of $\sim 67\%$, and in I-Love-Q universality by factors of $\sim50$\%, tabulated in Tab.~\ref{tab:maxVar}.
We also present new R-Love relations displaying the relationship between NS radius and tidal deformability, showing maximum EoS variability of $18.88\%$ from the fit discussed in Sec.~\ref{sec:clove}.
Additionally, we find that hybrid quark-hadron stars do obey the I-Love-Q and C-Love relations, with sightly increased EoS variability.
We also find that binaries consisting of one massive hybrid star, and one small-mass hadronic star do not agree with derived binary Love relations, due to the large separation in $\tilde{\Lambda}$ between the constituent stars.
Due to the current limitations in detector sensitivity, the systematic errors arising from EoS variation in the EoS-insensitive relations are far outweighed by the statistical errors accrued by extraction of high-order PN tidal parameters from GW waveforms.
Fig.~\ref{fig:stackedFisher} compiles the results from Fisher analyses approximating the uncertainties accrued from parameter extraction for GW170817 as if it had been observed by future detectors $A \equiv ($ O2~\cite{aLIGO}, aLIGO~\cite{aLIGO}, A\texttt{+}~\cite{Ap_Voyager_CE}, Voyager~\cite{Ap_Voyager_CE}, ET~\cite{ET}, CE~\cite{Ap_Voyager_CE}$)$.
Here, $\sigma^A_{\text{GW170817}}$ and $\rho^A_{\text{GW170817}}$ correspond to the statistical error and signal-to-noise-ratio found from observing GW170817 on detector $A$, and $\sigma^A_N$ represents the combined uncertainty after the detection of $N_A$ events (corresponding to the binary NS merger rate associated interferometer $A$).
These values are further tabulated in Table~\ref{tab:variances}.
It can be seen from this figure that the statistical errors from parameter extraction become comparable the EoS-insensitive relation systematics (indicated by the magenta horizontal line) for the Voyager telescope - indicating when improved binary Love relations will become necessary.
Also shown in the figure is the same measurement uncertainties upon the injection of the PhenomD + NRTidal waveform, rather than the PhenomD + 6PN tidal correction to demonstrate the effect of waveform mismodeling.
\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{stackedFisher.eps}
\end{center}
\caption{
Estimated statistical uncertainties $\sigma^A_{\text{GW170817}}$ (blue circle) in the extraction of $\lambda_0$ from GW170817 (O2) as if observed with future interferometers aLIGO, A\texttt{+}, Voyager, CE, and ET-D as a function of the signal-to-noise-ratio $\rho^A_{\text{GW170817}}$.
This is compared to the combined uncertainty $\sigma^A_N$ (turquoise shaded region) from the observation of multiple events from a simulated population of size $N_A$ corresponding to the upper and lower limits of expected binary NS merger detection rates for O2, aLIGO, A\texttt{+}, Voyager, CE, and ET-D over 1 observation year.
It can be seen that the statistical errors in parameter extraction become comparable to the systematics from EoS variation in binary Love relations (purple shaded region) for Voyager and on.
Note there is no simulated populations of events under the O2 design sensitivity - due to there only being a single event during the observing period.
These results are further tabulated in Table~\ref{tab:variances}.
Additionally shown in the figure are the estimated statistical uncertainties on the extraction of $\lambda_0$ upon the injection of the PhenomD + NRTidal correction waveform (IMRD + NRTidal), rather than the PhenomD + 6PN tidal correction (IMRD + 6PN), in order to demonstrate the waveform mismodeling uncertainties.
The single-detection uncertainties are shown in unfilled blue circles, while the combined detection uncertainties are outlined in dashed maroon.
}
\label{fig:stackedFisher}
\end{figure}

%outline
The organization of this paper is outlined below.
We begin with a complementary background and theory material in Sec.~\ref{sec:theory}.
We continue in Sec.~\ref{sec:universal} by finding new and improved binary Love, I-Love-Q, and C-Love relations, and considering how well hybrid star EoSs agree with the EoS-insensitive relations.
We next examine these improved EoS-insensitive relations and question whether or not they are useful for future interferometers in Sec.~\ref{sec:observations}.
We conclude in Sec.~\ref{sec:conclusion} by discussing our results and mentioning avenues of future work.
Throughout this paper, we have adopted geometric units of $G=c=1$, unless otherwise stated.

@@

\subsection{Spectral representations of neutron star equations of state}\label{sec:eos}
The structure of a NS and its tidal interactions in a binary system rely heavily on the underlying state function describing the relationship between pressure ($p$) and energy density ($\epsilon$), or equation of state, of nuclear matter.
Given that all currently proposed EoSs utilize certain approximations~\cite{Oertel:Review,Baym:Review}, one method to study a wide range of physically realizable EoSs is to parameterize them such that any realistic EoS can be represented with a small number of parameters.
Spectral representations~\cite{Lindblom:2010bb,Lindblom:2012zi,Lindblom:2013kra,Lindblom:2018rfr,Abbott:2018exr} parameterize EoSs by performing spectral expansions on the adiabatic index $\Gamma(p)$\footnote{Another way of parameterizing EoSs is the piecewise polytropic formulation~\cite{Read2009,Lackey:2014fwa,Carney:2018sdv}.}:
\begin{equation}
\Gamma(x) = \exp{\sum_k\gamma_k x^k},
\end{equation}
where $x \equiv \log{p/p_0}$ for minimum pressure $p_0$.
The equation of state is then determined by an integration of the differential equation:
\begin{equation}
\frac{d \epsilon(p)}{dp}=\frac{\epsilon(p)+p}{p \Gamma(p)}.
\end{equation}
Using this formalism, any valid EoS can be approximated through the choice of 4 or more spectral coefficients $\gamma_k$, tabulated for several common EoSs in Table 1 of~\cite{Lindblom:2018rfr}.

@@

In addition, we investigate 10 transitional quark-hadron matter stars, which undergo first-order phase transitions at pressure $P_{\text{tr}}$, where the hadronic branch departs into a quark-matter branch at a given mass transition~\cite{Paschalidis2018,Alford:2017qgh,1971SvA....15..347S,Zdunik:2012dj,Alford:2013aca}.
In particular, we focus on the ACS and ACB models described in Ref.~\cite{Paschalidis2018}, shown in Fig.~\ref{fig:hybridML}.
These result in two distinct types of NSs, based on their observed mass: (i) massive ($M \geq \text{ M}_{\text{tr}}$) hybrid stars which have quark-matter cores and nuclear matter crusts (we denote this as ``HS"), and (ii) small-mass ($M \leq \text{ M}_{\text{tr}}$) hadronic stars with no internal transition to quark matter (we denote this as ``NS").

\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{EoSs.eps}
\end{center}
\caption{
Small representative sets of the unconstrained (dotted) and constrained (solid) EoSs used in this analysis. 
The constrained EoSs were populated by randomly selecting 100 posterior samples obtained from the GW170817 90\% credible level shown in cyan~\cite{LIGO:posterior}.
Additionally, 100 unconstrained EoSs were generated by randomly sampling spectral coefficients of physically valid EoSs, with no further constrained applied.
The following constraints were also applied to both sets of EoSs: (i) the EoS must have a causal structure within 10\%, and (ii) the EoS must support maximum NS mass of at least $1.97 \text{ M}_{\odot}$, consistent with astrophysical observations.
Finally, the high density core EoSs generated above were stitched to the low-density crust EoS of SLy~\cite{Douchin:2001sv} at $\rho_{\text{stitch}}=1.3 \times 10^{14} \text{ g/cm}^3$.
}
\label{fig:eos}
\end{figure}

@@

Next, we consider the case of two NSs in a binary, such as the system found in GW170817, where each star 
individually experiences a neighboring tidal field.
Thus each star possesses tidal deformabilities $\Lambda_1$ and $\Lambda_2$ entering the gravitational waveform.
Due to strong correlations between these parameters, individual extraction is very difficult with current interferometer sensitivity limitations.
However, the waveform templates can be strategically reparameterized to instead include independent linear combinations of $\Lambda_1$ and $\Lambda_2$ to mitigate correlations between them. 
Any set of two independent functions may be used, for example: tidal parameters $\tilde{\Lambda}=\tilde{\Lambda}(\Lambda_1,\Lambda_2)$ and $\delta \tilde{\Lambda}=\delta \tilde{\Lambda}(\Lambda_1,\Lambda_2)$~\cite{Wade:tidalCorrections} first enter the gravitational waveform at 5PN and 6PN orders respectively, thus partially breaking degeneracies.
Here, $\tilde{\Lambda}$ is known as the mass weighted tidal deformability, or \emph{chirp deformability} due to its status as the dominant tidal parameter in the waveform.

\subsection{EoS insensitive relations}\label{sec:eosInsensitive}
Current gravitational wave interferometry is not yet sensitive enough to accurately extract both tidal parameters $\tilde{\Lambda}$ and $\delta\tilde{\Lambda}$.
In a search to remedy this, Yagi and Yunes~\cite{Yagi:binLove} found that symmetric and antisymmetric combinations of tidal deformabilities:
\begin{equation}
\Lambda_s \equiv \frac{\Lambda_1 + \Lambda_2}{2}, \hspace{6mm} \Lambda_a \equiv \frac{\Lambda_1 - \Lambda_2}{2},
\end{equation}
display EoS-insensitive properties to a high degree, showing EoS variations up to 20\% for binaries with masses less than $1.7 \text{ M}_{\odot}$, for a representative set of 11 EoSs. 
These ``binary Love relations" allow one to break degeneracies between coupled tidal parameters.
This is important for two reasons: (i) allows us to algebraically eliminate tidal parameters from the template parameter list, improving parameter estimation, and (ii) automatic measurement of second tidal parameter given observation of the first.
It was shown through a simple Fisher analysis that the use of binary Love relations improved parameter extraction on $\tilde{\Lambda}$ by an order of magnitude.

Similar EoS-insensitive relations have been found to exist between individual NS observables: moment of inertia (I), tidal deformability (Love), quadrupole moment (Q), and compactness (C), known as the ``I-Love-Q" and ``C-Love" relations~\cite{Yagi:ILQ, Yagi:binLove}.
These have been found to be EoS insensitive by up to 1\% and 6.5\% respectively.
In this paper, we show improvement in both of these EoS-insensitive relations by restricting to EoSs sampled from GW170817s posterior on pressure as a function of density.
%%%%%%%%%%%%%%%%%%%%%%%%% Begin universal relations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{EoS-insensitive relations}\label{sec:universal}
In this section, we follow the analyses performed in~\cite{Yagi:binLove,Yagi:ILQ} with two new sets of 100 spectrally generated EoS: (i) posterior samples randomly selected from GW170817s 90\% posterior on on pressure as a function of density as shown by the solid green curves in Fig.~\ref{fig:eos}, and (ii) those unconstrained by any prior EoS information, as shown by the dotted maroon curves in Fig.~\ref{fig:eos}.

\subsection{I-Love-Q relations}\label{sec:ilq}
Here we present our results on I-Love-Q universality as compared to Fig. 1 of Yagi and Yunes~\cite{Yagi:ILQ}.
In particular, we consider two distinct classes of NSs: nuclear matter EoSs and hybrid quark-hadron star EoSs as described in Sec.~\ref{sec:theory}.
We begin in Sec.~\ref{sec:ilq-nuc} by fitting the new I-Love-Q relations using the constrained set of EoSs.
This is followed in Sec.~\ref{sec:ilq-hyb} by an analysis and discussion into how well hybrid stars agree with the improved binary Love relations.

{\ny{Overall, let's not call these relations ``I-Love-Q relations'' or ``binary love relations.''  It's much better to be brief and simply call them I-Love-Q relations and Binary Love relations (like is done in the literature). It should be clear when we define them that they are EOS insensitive}}

>>>>

{\ny{Just some general comments so that we don't forget:
\begin{itemize}
\item The EoS-insensitive relations are called ``binary Love'', ``I-Love-Q'', ``C-Love'' and ``R-Love.'' There is no need to call them ``binary Love EoS-insensitive relations.'' That's too verbose and not what they are called in the literature. I am ok with describing these relations as EoS-insensitive instead of approximately universal, but let's not change the names. 
\item There is some serious tidying up that needs to be done to all of the text. Specifically, there are missing prepositions, like ``the'', in lots of places that makes the reading a bit choppy and we can improve. Zack, when you have a chance, please do a detail read through and fix the flow as much as you can. 
\item Eventually, we need to discuss the abstract and the intro in more detail. 
\item Let's not forget to check the acknowledgments for funding sources. 
\item Let's do a spell check. 
\item Let's remove all commented out text.  
\end{itemize}
}}
The dependence of pressure on density for cold supranuclear matter found inside neutron stars (NSs) remains to be one of the largest mysteries in both nuclear physics and astrophysics.
This internal structure of such stars, known as the equation of state (EoS), is exceedingly important, as it is the determining factor for many NS observables - such as the mass, the radius, and many more.
Unfortunately, terrestrial experiments can only study the EoS to around the nuclear saturation density ($\rho_{\text{sat}} \approx 2.5 \times 10^{14} \text{ g/cm}^3$)~\cite{Li:HeavyIon,Tsang:SymmetryEnergy,Centelles:NeutronSkin,Li:CrossSections,Chen:SymEnergy} (although some temperature-dependent heavy-ion collision experiments may probe higher densities~\cite{Abrahamyan:2012gp,Horowitz:2012tj,Horowitz:2013wha}), making NSs ideal laboratories for constraining ultra dense nuclear matter.

@@

%Universal relations
Unfortunately, current detectors are only sensitive enough to accurately measure the dominant tidal parameter, $\lambda_0$.
Previous work by Yagi and Yunes~\cite{Yagi:binLove} resolved this issue by finding approximately EoS-insensitive ``binary Love relations" between symmetric and anti-symmetric combinations of tidal deformabilities $\Lambda_{s,a}=\frac{1}{2}(\Lambda_1 \pm \Lambda_2)$.
This allows one to further break degeneracies between tidal parameters, resulting in both (i) a decrease in uncertainty upon parameter extraction of $\Lambda_s$ on the removal of parameter $\Lambda_a$ from the model, and (ii) the ability to measure the secondary tidal parameter upon measurement of the first~\cite{Katerina:residuals}.  
Use of these relations leads to up to an order of magnitude improvement in parameter estimation of $\Lambda_s$ through a simple Fisher analysis~\cite{Yagi:binLove}, also confirmed with the detailed analysis of~\cite{Katerina:residuals}.
Similar ``I-Love-Q" and ``C-Love" relations~\cite{Yagi:ILQ} have been found between individual NS observables: quadrupole moment , moment of inertia, tidal deformability, and compactness; similarly allowing the removal of tidal parameters from the model list, and allowing an automatic determination of one observable through the measurement of another.
However, these relations have been tuned to a collection of EoS models, and each have intrinsic uncertainties associated with their use.

%How we improved this
In this analysis, we aim to improve this important work by imposing the restrictions found on the EoS~\cite{LIGO:posterior} derived from the recent binary NS merger observation, GW170817~\cite{TheLIGOScientific:2017qsa}.
By restricting to only EoSs which lie within the the 90\% credible region on pressure as a function of density found in~\cite{LIGO:posterior}, we repeat the analyses done in~\cite{Yagi:binLove,Yagi:ILQ} and show an increase in universality for I-Love-Q, C-Love, and binary Love relations.
We also investigate the relationship between the NS radius and its tidal deformability, or the R-Love relations, determining if the uncertainties can be reduced.
In addition, we consider hybrid star EoSs seen in~\cite{Paschalidis2018}, which experience strong first-order transitions from hadronic matter to quark matter.
These provide a departure from the nuclear matter relations considered here - providing valuable insight into them.
Further, we estimate the viability of using improved EoS-insensitive relations on future GW detections by performing simple Fisher analyses to approximate when the statistical errors from parameter extraction become comparable to the systematic errors due to EoS variation in the relations.
To estimate the effect of waveform mismodeling, we additionally consider two different tidal corrections to the injected PhenomD~\cite{PhenomDI,PhenomDII} waveform: 6PN~\cite{Wade:tidalCorrections} and NRTidal~\cite{Samajdar:NRTidal}

\subsection{Executive Summary}
\zc{Me and Kent discussed whether or not we should include this section or not, and we decided like having the quick summary for busy readers - especially with the increasing number of figs and tables (8 tables, 14 figures so far). We can discuss this further if anyone doesn't want it included :)}
{\ny{I'm ok with keeping this, but let's get the science finished first.}}
%What we did
In this paper we attempt to improve I-Love-Q, C-Love, and binary Love relations by imposing constraints on the EoS found by GW170817~\cite{LIGO:posterior,TheLIGOScientific:2017qsa}.
First, we generate two large samples of spectral EoSs~\cite{Lindblom:2018rfr}, where we (i) impose the restriction that they must be drawn from GW170817s 90\% credible posterior in pressure as a function of density, and (ii) don't impose any restrictions for comparison purposes.
Next, we follow the important works of~\cite{Yagi:binLove} and~\cite{Yagi:ILQ} to show how the ``constrained" set of EoSs show improved universality (binary Love, I-Love-Q, and C-Love) from both the previous works, and the ``unconstrained" sample, shown by Fig.~\ref{fig:ILQ}.
In addition, we determine the degree to which hybrid star EoSs obey binary Love relations.
Finally, through a simple Fisher analysis, we consider the value in improving EoS-insensitive relations by approximating when the statistical errors from parameter extraction become comparable to the systematic errors due to EoS variation in the relations.

%Final Results
Upon use of the constrained set of EoSs, we find an improvement in binary Love universality by a factor of $\sim 59$\% for stars with mass ratios of $0.75$, in C-Love universality by a factor of $\sim 67\%$, and in I-Love-Q universality by factors of $\sim50$\%, tabulated in Tab.~\ref{tab:maxVar}.
We also present new R-Love relations displaying the relationship between NS radius and tidal deformability, showing maximum EoS variability of $18.88\%$ from the fit discussed in Sec.~\ref{sec:clove}.
Additionally, we find that hybrid quark-hadron stars do obey the I-Love-Q and C-Love relations, with sightly increased EoS variability.
We also find that binaries consisting of one massive hybrid star, and one small-mass hadronic star do not agree with derived binary Love relations, due to the large separation in $\tilde{\Lambda}$ between the constituent stars.
Due to the current limitations in detector sensitivity, the systematic errors arising from EoS variation in the EoS-insensitive relations are far outweighed by the statistical errors accrued by extraction of high-order PN tidal parameters from GW waveforms.
Fig.~\ref{fig:stackedFisher} compiles the results from Fisher analyses approximating the uncertainties accrued from parameter extraction for GW170817 as if it had been observed by future detectors $A \equiv ($ O2~\cite{aLIGO}, aLIGO~\cite{aLIGO}, A\texttt{+}~\cite{Ap_Voyager_CE}, Voyager~\cite{Ap_Voyager_CE}, ET~\cite{ET}, CE~\cite{Ap_Voyager_CE}$)$.
Here, $\sigma^A_{\text{GW170817}}$ and $\rho^A_{\text{GW170817}}$ correspond to the statistical error and signal-to-noise-ratio found from observing GW170817 on detector $A$, and $\sigma^A_N$ represents the combined uncertainty after the detection of $N_A$ events (corresponding to the binary NS merger rate associated interferometer $A$).
These values are further tabulated in Table~\ref{tab:variances}.
It can be seen from this figure that the statistical errors from parameter extraction become comparable the EoS-insensitive relation systematics (indicated by the magenta horizontal line) for the Voyager telescope - indicating when improved binary Love relations will become necessary.
Also shown in the figure is the same measurement uncertainties upon the injection of the PhenomD + NRTidal waveform, rather than the PhenomD + 6PN tidal correction to demonstrate the effect of waveform mismodeling.
\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{stackedFisher.eps}
\end{center}
\caption{
Estimated statistical uncertainties $\sigma^A_{\text{GW170817}}$ (blue circles) in the extraction of $\lambda_0$ from GW170817 (O2 detector sensitivity) as if it were observed with future interferometers aLIGO, A\texttt{+}, Voyager, CE, and ET-D as a function of the signal-to-noise-ratio $\rho^A_{\text{GW170817}}$.
This is compared to the combined uncertainty $\sigma^A_N$ (turquoise shaded region) from the observation of multiple events from a simulated population of size $N_A$ corresponding to the upper and lower limits of the expected binary NS merger detection rates for O2, aLIGO, A\texttt{+}, Voyager, CE, and ET-D over 1 observation year.
It can be seen that the statistical errors in the parameter extraction become comparable to the systematics from EoS variation in improved binary Love relations (indigo dashed line) for Voyager CE, and ET.
These results are further tabulated in Table~\ref{tab:variances}.
Additionally shown in the figure are the estimated statistical uncertainties on the extraction of $\lambda_0$ upon the injection of the PhenomD + NRTidal correction waveform (IMRD + NRTidal), rather than the PhenomD + 6PN tidal correction (IMRD + 6PN), in order to demonstrate the effect of waveform mismodeling uncertainties.
The single-detection uncertainties are shown by unfilled blue circles, while the combined detection uncertainties are outlined by dashed maroon.
}
\label{fig:stackedFisher}
\end{figure}

%outline
The organization of this paper is outlined below.
We begin with a complementary background and theory material in Sec.~\ref{sec:theory}.
We continue in Sec.~\ref{sec:universal} by finding new and improved binary Love, I-Love-Q, and C-Love relations, and considering how well hybrid star EoSs agree with the EoS-insensitive relations.
We next examine these improved relations and question whether or not they are useful for future interferometers in Sec.~\ref{sec:observations}.
We conclude in Sec.~\ref{sec:conclusion} by discussing our results and mentioning avenues of future work.
Throughout this paper, we have adopted geometric units of $G=c=1$, unless otherwise stated.

@@

\subsection{Spectral representations of neutron star equations of state}\label{sec:eos}
The structure of a NS and its tidal interactions in a binary system rely heavily on the underlying state function (or equation of state) describing the relationship between pressure ($p$) and energy density ($\epsilon$) of nuclear matter.
Given that all currently proposed EoSs utilize certain approximations~\cite{Oertel:Review,Baym:Review}, one method to study a wide range of physically realizable EoSs is to parameterize them such that any realistic EoS can be represented with a small number of parameters.
Spectral representations~\cite{Lindblom:2010bb,Lindblom:2012zi,Lindblom:2013kra,Lindblom:2018rfr,Abbott:2018exr} parameterize EoSs by performing spectral expansions on the adiabatic index $\Gamma(p)$\footnote{Another way of parameterizing EoSs is the piecewise polytropic formulation~\cite{Read2009,Lackey:2014fwa,Carney:2018sdv}.}:
\begin{equation}
\Gamma(x) = \exp{\sum_k\gamma_k x^k},
\end{equation}
where $x \equiv \log{p/p_0}$ for a minimum pressure $p_0$.
The equation of state is then determined by an integration of the differential equation:
\begin{equation}
\frac{d \epsilon(p)}{dp}=\frac{\epsilon(p)+p}{p \Gamma(p)}.
\end{equation}
Using this formalism, any valid EoS can be approximated through the choice of 4 or more spectral coefficients $\gamma_k$, tabulated for several common EoSs in Table 1 of~\cite{Lindblom:2018rfr}.

@@

In addition, we investigate 10 transitional quark-hadron matter stars, which undergo first-order phase transitions at pressure $P_{\text{tr}}$, where the hadronic branch departs into a quark-matter branch at a given mass transition~\cite{Paschalidis2018,Alford:2017qgh,1971SvA....15..347S,Zdunik:2012dj,Alford:2013aca}.
In particular, we focus on the ACS and ACB models described in Ref.~\cite{Paschalidis2018}, shown in Fig.~\ref{fig:hybridML}.
These result in two distinct types of NSs, based on their observed mass: (i) massive ($M \geq \text{ M}_{\text{tr}}$) hybrid stars which have quark-matter cores and nuclear matter crusts (we denote this as ``HS"), and (ii) low-mass ($M \leq \text{ M}_{\text{tr}}$) hadronic stars with no internal transition to quark matter (we denote this as ``NS").

\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{EoSs.eps}
\end{center}
\caption{
Small representative sets of the unconstrained (dotted) and constrained (solid) EoSs used in this analysis. 
The constrained EoSs were populated by randomly selecting 100 posterior samples obtained from the GW170817 90\% credible level shown in cyan~\cite{LIGO:posterior}.
Additionally, 100 unconstrained EoSs were generated by randomly sampling spectral coefficients of physically valid EoSs, with no further constraints applied.
The following restrictions were also applied to both sets of EoSs: (i) the EoS must have a causal structure within 10\%, and (ii) the EoS must support maximum NS mass of at least $1.97 \text{ M}_{\odot}$, consistent with astrophysical observations.
Finally, the high density core EoSs generated above were stitched to the low-density crust EoS of SLy~\cite{Douchin:2001sv} at $\rho_{\text{stitch}}=1.3 \times 10^{14} \text{ g/cm}^3$.
}
\label{fig:eos}
\end{figure}

@@

Next we consider binary NS systems, as was found in GW170817, where each star individually experiences a neighboring tidal field.
Thus each star possesses tidal deformabilities $\Lambda_1$ and $\Lambda_2$ entering the gravitational waveform.
Due to strong correlations between these parameters, individual extraction is very difficult with current interferometer sensitivity limitations.
However, the waveform templates can be strategically reparameterized to instead include independent linear combinations of $\Lambda_1$ and $\Lambda_2$ to mitigate correlations between them. 
Any set of two independent functions may be used, for example: tidal parameters $\tilde{\Lambda}=\tilde{\Lambda}(\Lambda_1,\Lambda_2)$ and $\delta \tilde{\Lambda}=\delta \tilde{\Lambda}(\Lambda_1,\Lambda_2)$~\cite{Wade:tidalCorrections} first enter the gravitational waveform at 5PN and 6PN orders respectively, thus partially breaking degeneracies.
Here, $\tilde{\Lambda}$ is known as the mass weighted tidal deformability and is the dominant tidal parameter in the waveform.

\subsection{EoS insensitive relations}\label{sec:eosInsensitive}
Current gravitational wave interferometry is not yet sensitive enough to accurately extract both tidal parameters $\tilde{\Lambda}$ and $\delta\tilde{\Lambda}$.
In a search to remedy this, Yagi and Yunes~\cite{Yagi:binLove} found that symmetric and antisymmetric combinations of tidal deformabilities:
\begin{equation}
\Lambda_s \equiv \frac{\Lambda_1 + \Lambda_2}{2}, \hspace{6mm} \Lambda_a \equiv \frac{\Lambda_1 - \Lambda_2}{2},
\end{equation}
display EoS-insensitive properties to a high degree, showing EoS variations up to 20\% for binaries with masses less than $1.7 \text{ M}_{\odot}$, for a representative set of 11 EoSs. 
These ``binary Love relations" allow one to break degeneracies between coupled tidal parameters.
This is important for two reasons: (i) they allow one to algebraically eliminate tidal parameters from the template parameter list, improving overall parameter estimation, and (ii) they allow the automatic measurement of a second tidal parameter given observation of the first.
It was shown through a simple Fisher analysis that the use of binary Love relations improved parameter extraction on $\tilde{\Lambda}$ by up to an order of magnitude.

Similar EoS-insensitive relations have been found to exist between individual NS observables: moment of inertia (I), tidal deformability (Love), quadrupole moment (Q), and compactness (C), known as the ``I-Love-Q" and ``C-Love" relations~\cite{Yagi:ILQ, Yagi:binLove}.
These have been found to be EoS insensitive by up to 1\% and 6.5\% respectively.
In this paper, we show improvement in all of these relations by restricting to EoSs sampled from GW170817s posterior on pressure as a function of density.
%%%%%%%%%%%%%%%%%%%%%%%%% Begin universal relations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{EoS-insensitive relations}\label{sec:universal}
In this section, we follow the analyses performed in~\cite{Yagi:binLove,Yagi:ILQ} with two new sets of 100 spectrally generated EoS: (i) posterior samples randomly selected from GW170817s 90\% posterior on on pressure as a function of density as shown by the solid green curves in Fig.~\ref{fig:eos}, and (ii) those unconstrained by any prior EoS information, as shown by the dotted maroon curves in Fig.~\ref{fig:eos}.

\subsection{I-Love-Q relations}\label{sec:ilq}
Here we present our results on I-Love-Q universality as compared to Fig. 1 of Ref.~\cite{Yagi:ILQ}.
In particular, we consider two distinct classes of NSs: nuclear matter EoSs and hybrid quark-hadron star EoSs as described in Sec.~\ref{sec:theory}.
We begin in Sec.~\ref{sec:ilq-nuc} by fitting the new I-Love-Q relations using the constrained set of EoSs.
This is followed in Sec.~\ref{sec:ilq-hyb} by an analysis and discussion into how well hybrid stars agree with the improved binary Love relations.
